---
title: "p8105_hw3_pm3148"
author: "Pooja Mukund"
date: "10/15/2021"
output: github_document
---

# Load Libraries
```{r}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
```
# Problem 1: Instacart
```{r}
#Load data
data("instacart")
```

```{r}
#How many aisles are there, and which aisles are the most items ordered from?
instacart%>%
  group_by(aisle)%>%
  summarize(count = n())%>%
  arrange(desc(count))
```

The table shows that we have 134 unique aisles represented in the dataset. Fresh vegetables, fresh fruits, and packages vegetable fruits are the aisles with the largest amount of records/purchases. 


```{r}
#Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
instacart%>%
  group_by(aisle)%>%
  summarize(count = n())%>%
  filter(count > 10000)%>%
  arrange(desc(count))%>%
  ggplot(aes(x=reorder(aisle, count), y = count)) + 
  geom_bar(stat="identity")+
  coord_flip()+
  labs(caption = "Limited to aisles with more than 10000 items ordered") + 
  xlab("Aisle Name") + ylab("Count of Items Ordered") + ggtitle("Number of items ordered by aisle") +
  theme(axis.text.y = element_text(size =7))

```

Fresh vegetables, fresh fruits, and packaged vegetables fruits have the highest amount of items ordered from. Butter, oils vinegars, and dry pasta have the lowest amount of items in this subset (items > 10,000). 
```{r}
#Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
instacart%>%
  group_by(aisle, product_name)%>%
  filter(aisle == "baking ingredients" | aisle =="dog food care" | aisle =="packaged vegetables fruits")%>%
  summarise(count = n())%>%
  top_n(n=3, wt=count)%>%
  arrange(desc(count), .by_group = TRUE)

instacart%>%
  group_by(aisle, product_name)%>%
  filter(aisle == "baking ingredients" | aisle =="dog food care" | aisle =="packaged vegetables fruits")%>%
  summarise(count = n())%>%
  filter(min_rank(desc(count))<4)%>%
  arrange(desc(count), .by_group = TRUE)
  
  
```

The top three most popular items in the baking ingredients aisles are: Light Brown Sugar, Pure Baking Soda, and Cane Sugar. In dog food care, the three most popular items are Snack Sticks Chicken & Rice Recipe, Organix Chicken and Brown Rice Recipe, and Small Dog Biscuits. The most popular items in packaged vegetable fruits are organic baby spinach, organic raspberries, and organic blueberries. The packaged vegetables fruits aisle category has the highest volume of products being sold with 9784 baby spinach, 5546 raspberries, and 4966 blueberries. Baking ingredients had a volume in the 100s and dog food care had volume between 26-30. 
```{r}
#Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
  instacart%>%
  group_by(product_name, order_dow)%>%
  filter(product_name =="Pink Lady Apples" | product_name =="Coffee Ice Cream")%>%
  summarise(mean_hour = mean(order_hour_of_day))%>%
  pivot_wider(names_from = order_dow,
              values_from = mean_hour)%>%
  rename(Sunday =`0`,
         Monday =`1`,
         Tuesday =`2`,
         Wednesday =`3`,
         Thursday =`4`,
         Friday =`5`,
         Saturday =`6`)
```

Coffee Ice Cream was ordered the most on Thursdays. Pink Lady Apples were ordered the most on Wednesdays. Coffee Ice Cream is ordered more than Pink Lady Apples on most days except for Friday. 

# Problem 2 
```{r}
#Data Cleaning 
data("brfss_smart2010")

brfss<- brfss_smart2010%>%
  janitor::clean_names()

brfss%>%
  distinct(topic)
#Overall Health topic is all coded the same way ("Overall Health")

brfss%>%
  filter(topic == "Overall Health")%>%
  distinct(response)
#All responses are labelled in the same format

#Change response to factor variable 
brfss_cln<- brfss%>%
  filter(topic == "Overall Health")%>%
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))

```
```{r}
#In 2002, which states were observed at 7 or more locations? What about in 2010?
brfss_cln%>%
  filter(year==2002)%>%
  group_by(locationabbr)%>%
  summarise(location_count = n_distinct(locationdesc))%>%
  filter(location_count >=7)

#In 2010 
brfss_cln%>%
  filter(year==2010)%>%
  group_by(locationabbr)%>%
  summarise(location_count = n_distinct(locationdesc))%>%
  filter(location_count >=7)
```   

In 2002, there were 6 states observed at 7 or more locations: CT, FL, MA, NC, NJ, and PA. In 2010, there were 14 states observed at 7 or more locations: CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA. FL, MA, NC, NJ, and PA were observed at 7 or more locations in both years.  
```{r}
##Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

brfss_cln%>%
 filter(response =="Excellent")%>%
  select(year, locationabbr, data_value)%>%
  group_by(year, locationabbr)%>%
  mutate(state_avg = mean(data_value, na.rm=TRUE))%>%
  ggplot(aes(x=year, y = state_avg, col = locationabbr)) +
  geom_line() +
  xlab('Year') + 
  ylab('State Average Value')+
  ggtitle("Average Value across Year by State")+
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

  
```

Plot is showing all of the state averages over time in years. No obvious trends in this plot. There is a lot of variability in average value year to year for each state. Not the most informative plot. It might be better to separate by region (ex. Northeast, Southeast, etc.) so that you can pick up on distinct trends. 
```{r}
# Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
brfss_cln%>%
  filter(locationabbr=="NY")%>%
  filter(year ==2006 | year==2010)%>%
  ggplot(aes(x = data_value)) +geom_histogram() +facet_wrap(year~response, nrow=2)

brfss_cln%>%
  filter(locationabbr=="NY")%>%
  filter(year ==2006 | year==2010)%>%
  ggplot(aes(x = data_value, y = response))+ 
  geom_density_ridges(scale = 0.85) +
  facet_grid(rows=vars(year))+
  xlab('Data Value')+
  ylab('Response')+
  ggtitle("Distribution of Data Value for Year 2006 and 2010")
```   

Plot shows that there is not much of a discrepancy in data values between years 2010 and 2006. Responses for Good, Very Good, and Excellent had generally higher data_value scores than Fair and Poor. 

## Problem 3
```{r}
#Load accelorometer data
accel<- read_csv('./data/accel_data.csv')

#Add weekend variable 
accel<- accel%>%
  janitor::clean_names()%>%
  mutate(weekend = ifelse(day =="Saturday" | day =="Sunday", "weekend", "weekday"), 
         day = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday","Friday", "Saturday")))



```

Resulting dataset has 35 variables and 1444 observations. The variables include week, day id, day, activity data from minute 1 to minute 1440 of the 24 hour period, and the weekend variable which classifies the observation as a weekend or weekday. 

```{r}
#Total Activity Variable per Day of Week 
accel%>%
  mutate(activity_sum_min = (accel%>%
                               select(starts_with("activity"))%>%
                               rowSums()))%>%
  select(day, activity_sum_min)%>%
  group_by(day)%>%
  summarise(activity_sum = sum(activity_sum_min))%>%
knitr::kable()

#Total Activity per Day 
accel%>%
  mutate(activity_sum_min = (accel%>%
                               select(starts_with("activity"))%>%
                               rowSums()))%>%
  select(day_id, activity_sum_min)

accel%>%
  mutate(activity_sum_min = (accel%>%
                               select(starts_with("activity"))%>%
                               rowSums()))%>%
  select(day_id, activity_sum_min)%>%
  ggplot(aes(x = day_id, y = activity_sum_min)) + 
  geom_point()+
  geom_line()+
  geom_smooth()+
  xlab("Day")+
  ylab("Activity Sum")
```
Looking at the the table for activity per Day of Week, we can see that Friday has the highest activity and then Wednesday followed by Thursday. Saturday has the lowest activity. 

There are certain days with only 1440 activity recorded. These may be days where accelerometer did not accurately pick up movement/activity data. I have included a plot of the data values by day ID to have a visual representation. It seems that in the first 15 days there is increasing activity and after day 15 there is a downward trend which picks back up around day 25. 

```{r}
#Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
accel%>%
  pivot_longer(activity_1:activity_1440,
               names_to="activity", 
               values_to = "activity_counts")%>%
  separate(activity, c(NA, "minute_of_day"))%>%
  mutate(minute_of_day = as.numeric(minute_of_day))%>%
  ggplot(aes(x = minute_of_day, y = activity_counts, col=day))+
  geom_line() +
  xlab('Minute of Day')+
  ylab('Activity Count')+
  ggtitle("Activity Count per Minute")+
  viridis::scale_color_viridis(
    name = "Day of Week", 
    discrete = TRUE
  )

``` 

This plot shows the activity count by minute of day for each day in the study (Days 1-35). Color represents the day of the week. The plot shows that Wednesdays tend to have lower activity counts than other days of the week. Plot is a little messy, might be more informative if we facet_grid by day. 
